Topology of the $O(3)$ non-linear sigma model under the gradient flow

### 2021-04-29 final-presentation.latex/main.tex
add final presentation

\setbeamerfont{caption}{size=\scriptsize}

@@

\usepackage{graphicx} % Allows including images

\usepackage{subfigure}
\usepackage{booktabs} % Allows the use of \toprule, \midrule and \bottomrule in tables

%----------------------------------------------------------------------------------------
%	TITLE PAGE
%----------------------------------------------------------------------------------------

\title[Short title]{Defining condensates in low-dimensional spin models through the gradient flow} % The short title appears at the bottom of every slide, the full title is only on the title page

\author{Stuart Thomas} % Your name
\institute[W\&M] % Your institution as it will appear on the bottom of every slide, may be shorthand to save space
{
Christopher Monahan \\ % Your institution for the title page
\medskip
}
\date{November 16, 2020} % Date, can be changed to a custom date

@@

\begin{frame}
\titlepage % Print the title page as the first slide
\end{frame}

@@

\begin{frame}
\frametitle{Power-divergent mixing in lattice QCD}
\begin{itemize}
    \item On the lattice, some operators mix and lead to divergences in the continuum limit.
    \item Example: breaking of rotational symmetry
    \item Operators must have different mass dimension
    \item In QCD:
        \begin{itemize}
            \item occurs with twist-2 operators
            \item used to calculate parton distribution functions
        \end{itemize}
\end{itemize}
\end{frame}

\begin{frame}
    \frametitle{Possible Solution: Smeared Operator Product Expansion (sOPE) }
    \begin{itemize}
        \item Expand non-local operators in a basis of smeared operators.
        \item Takes the form of a Laurent series
    \end{itemize}
    \textbf{Smeared Operator}: locally averaged operator
\end{frame}

@@

        \begin{figure}[h]
  \centering
      \subfigure[$\tau=0$]{\includegraphics[width=0.20\textwidth]{imgs/0.png}}
      \subfigure[$\tau=0.5$]{\includegraphics[width=0.20\textwidth]{imgs/0_5.png}}
      \subfigure[$\tau=2$]{\includegraphics[width=0.20\textwidth]{imgs/2.png}} 
      \subfigure[$\tau=10$]{\includegraphics[width=0.20\textwidth]{imgs/10.png}}
          \caption{Effect of flow time evolution on a random lattice in the symmetric phase. Red and blue indicate positive and negative values of the field in 2D spacetime.}
  \label{fig:flow}
\end{figure}

@@

\begin{frame}
\frametitle{Non-Linear $\sigma$ Model}
\begin{itemize}
    \item a nonperturbative model in Quantum Field Theory which provides a good toy model for Quantum Chromodynamics
    \item characterized by the action
        \begin{equation*}
            S = \frac{1}{2g} \int d^2x \: \partial_\mu \vec{\phi}(x) \cdot \partial^\mu \vec{\phi}(x)
        \end{equation*}
    where the $\vec{\phi}(x)$ is a 3-component vector with norm 1 and $g$ is a coupling constant. 
\item Merits 
    \begin{itemize}
        \item mass gap
        \item asymptotic freedom
        \item $O(2)$ renormalizability
        \item in condensed matter, models Heisenberg ferromagnets
    \end{itemize}
\end{itemize}
\end{frame}

@@

    \end{enumerate}
    \begin{figure}[h]

      \centering
          \subfigure[before]{\includegraphics[width=0.20\textwidth]{imgs/wolff1}}
          \subfigure[after]{\includegraphics[width=0.20\textwidth]{imgs/wolff2}}
          \caption{Wolff cluster algorithm applied on $\phi^4$ lattice in $1+1$ dimensions, red corresponding to positive values and green to negative}
      \label{fig:flow}
    \end{figure}
\end{frame}

\begin{frame}
    \frametitle{Checkerboard Parallelization Algorithm}
    \begin{itemize}
        \item Split lattice into ``white'' and ``black'' sites. 
        \item Then, perform metropolis sweep on white sites in parallel, then black sites.
        \item Ensures independence of sweeps.
\end{itemize}

    \begin{figure}[h]

      \centering
          \includegraphics[width=0.5\textwidth]{imgs/checkerboard}
          \caption{Illustration of checkerboard algorithm\footnote{Yang, Kun, et al. "High performance Monte Carlo simulation of Ising model on TPU clusters." Proceedings of the International Conference for High Performance Computing, Networking, Storage and Analysis. 2019.}}
    \end{figure}

@@

\begin{frame}
\frametitle{Comparison between Python and C++ code}
\begin{figure}[h]
  \centering
      \includegraphics[width=0.9\textwidth]{imgs/compare.png}
      \caption{Absolute difference between Python and C++ results of average $\bar\phi$, Binder cumulant and susceptibility for various $m_0^2$ values. $\lambda=0.5$, $L=64$, $1000$ measurements, $1000$ sweep thermalization with measurements taken every $100$ sweeps. }

>>>>

%\setbeamerfont{caption}{size=\scriptsize}

@@

\usepackage{graphicx} % Allows including images
\usepackage{tikz}
\usepackage[duration=30,disablemarkdown]{pdfpc}
\usepackage[font=tiny, labelfont=bf]{caption}
\usepackage{subcaption}
\usepackage{booktabs} % Allows the use of \toprule, \midrule and \bottomrule in tables

\renewcommand{\note}{\pdfpcnote}
%----------------------------------------------------------------------------------------
%	TITLE PAGE
%----------------------------------------------------------------------------------------

\title[Short title]{Topology of the $O(3)$ non-linear sigma model under the gradient flow} % The short title appears at the bottom of every slide, the full title is only on the title page

\author{Stuart Thomas} % Your name
\institute[W\&M] % Your institution as it will appear on the bottom of every slide, may be shorthand to save space
{
Christopher Monahan \\ % Your institution for the title page
\medskip
}
\date{May 3, 2021} % Date, can be changed to a custom date

\newcommand{\e}{\vec e}
\renewcommand{\theequation}{\relax}

@@

\begin{frame}
\titlepage % Print the title page as the first slide
\note{ - Test note one \\
- Test note two%
}
\end{frame}

\begin{frame}
    \frametitle{Quantum Field Theory}
\begin{equation}
    \label{eq:pathintegral}
    \langle \hat O \rangle = \frac{1}{Z} \int \mathcal{D}\phi \: \hat O [\phi]\; e^{iS[\phi]}
\end{equation}
\end{frame}

\begin{frame}
\frametitle{$\phi^4$ model}
\begin{equation}
    \label{eq:phi4 euclidean action}
    S_E[\phi] = \int d^2 x_E \left[\frac{1}{2}\left(\partial_t \phi\right)^2 + \frac{1}{2} \left(\partial_x \phi \right)^2 + \frac{1}{2} m_0^2 \phi^2 + \frac{\lambda}{4}\phi^4\right].
\end{equation}
\end{frame}

\begin{frame}
\frametitle{Non-Linear $\sigma$ Model}
\begin{equation}
    \label{eq:nlsm euclidean action}
    % S_E = \frac{\beta}{2} \int d^2x \; \partial^\mu \e \cdot \partial_\mu \e
    S_E = \frac{\beta}{2} \int d^2x \; \left[ \left(\partial_t \e\, \right)^2+ \left( \partial_x \e\,\right)^2 \right]
\end{equation}
\begin{itemize}
    \item a nonperturbative model in Quantum Field Theory which provides a good toy model for Quantum Chromodynamics
    \item characterized by the action
        \begin{equation*}
            S = \frac{1}{2g} \int d^2x \: \partial_\mu \vec{\phi}(x) \cdot \partial^\mu \vec{\phi}(x)
        \end{equation*}
    where the $\vec{\phi}(x)$ is a 3-component vector with norm 1 and $g$ is a coupling constant. 
\item Merits 
    \begin{itemize}
        \item mass gap
        \item asymptotic freedom
        \item $O(2)$ renormalizability
        \item in condensed matter, models Heisenberg ferromagnets
    \end{itemize}
\end{itemize}
\end{frame}

\begin{frame}
\frametitle{Monte Carlo and the Markov Chain}
\end{frame}

\begin{frame}
\frametitle{Topological Charge}
\begin{figure}
    \includegraphics[width=0.3\paperwidth]{imgs/homotopy.png}
    \caption{\label{fig:homotopy} Homotopy group}
\end{figure}
\note{- Test note one \\ - Test note two}
\end{frame}

\begin{frame}
\frametitle{Topological Action}
\begin{equation}
    S[\e\,] \rightarrow S[\e\,] - i \theta Q[\e\,].
\end{equation}
\end{frame}

\begin{frame}
    \frametitle{Topology Visualization}
\begin{figure}[h]
    \begin{subfigure}[b]{0.4\paperwidth}
    \centering
    \begin{tikzpicture}[scale=0.5, every node/.style={scale=0.7}]
        \draw[step=4cm,gray,thin] (-2,-2) grid (6,6);
        \draw[dotted] (0,0) -- (4,4);
        \draw[thick](0,0) node[circle,fill,inner sep=0pt, minimum size=0.3cm]{} node[anchor=north west] {$x_1$} -- 
                    (0,4) node[circle,fill,inner sep=0pt, minimum size=0.3cm]{} node[anchor=north west] {$x_2$} -- 
                    (4,4) node[circle,fill,inner sep=0pt, minimum size=0.3cm]{} node[anchor=north west] {$x_3$} -- 
                    (4,0) node[circle,fill,inner sep=0pt, minimum size=0.3cm]{} node[anchor=north west] {$x_4$} -- cycle ;
                    \draw[] (2,2) node[circle,fill,inner sep=0pt, minimum size=0.3cm]{} node[anchor=north east]{$x^*$};

        \draw[-stealth] (0.2,1) -- (0.2,3);
        \draw[-stealth] (1,3.8) -- (3,3.8);
        \draw[-stealth] (3,3.5) -- (1,1.5);

        \draw[-stealth] (1,0.5) -- (3,2.5);
        \draw[-stealth] (3.8,3) -- (3.8,1);
        \draw[-stealth] (3,0.2) -- (1,0.2);
    \end{tikzpicture}
    %\caption{\label{fig:plaquette} Visualization of plaquette $x^*$. The dotted line separates the plaquette into two signed areas which are used to define the topological charge density $q(x^*)$. Arrows represent order of signed area.}
    \end{subfigure} %
    \begin{subfigure}[b]{0.4\paperwidth}
    \centering
    \begin{tikzpicture}[scale=0.7, every node/.style={scale=0.7}]
        \pgfdeclarelayer{nodelayer}
        \pgfdeclarelayer{edgelayer}
        \pgfdeclareradialshading{sphere4}{\pgfpoint{-0.2cm}{0.5cm}}% 
            {rgb(0cm)=(1,1,1);
            rgb(1cm)=(0.5,0.5,0.5); rgb(1.05cm)=(1,1,1)}
            %rgb(0.7cm)=(0.1,0.1,0.1); rgb(1cm)=(0.5,0.05,0); rgb(1.05cm)=(1,1,1)}
        \pgfsetlayers{nodelayer,edgelayer}
        \tikzstyle{label}=[fill=none, draw=none, shape=circle]
        \tikzstyle{point}=[inner sep=0pt, minimum size=0.2cm,fill=black, draw, shape=circle]
        %\tikzstyle{interior line}=[{Stealth[scale=1.5]}-,dotted,thick]
        \tikzstyle{interior line}=[dotted,thick]

        \tikzstyle{triangle}=[thick]
        \tikzstyle{arrow}=[->, thick]
        \begin{pgfonlayer}{nodelayer}
            \node [style=point] (4) at (0.5, 2.5) {};
            \node [style=point] (5) at (0.55, 0.625) {};
            \node [style=point] (6) at (2.25, 0.75) {};
            \node (7) at (0, 0) {};
            \node (8) at (1.25, 1.375) {};
            \node (9) at (2.425, 2.525) {};
        \end{pgfonlayer}
        \begin{pgfonlayer}{edgelayer}
            \draw (0,0) circle (3cm);
            %\shade[inner color=white,outer color=lightgray] (0,0) circle (3cm);
            \shade[shading=sphere4] (0,0) circle (3cm);
            \draw [bend left=15,style=triangle] (4.center) to (6.center);
            \draw [bend left=355,style=triangle] (6.center) to (5.center);
            \draw [bend right=5,style=triangle] (5.center) to (4.center);
            \draw [style=interior line] (4.center) to (7.center);
            \draw [style=interior line] (5.center) to (7.center);
            \draw [style=interior line] (6.center) to (7.center);
            \draw [fill=gray!80] (4.center) to [bend left=15] (6.center) to [bend left=355] (5.center) to [bend right=5] cycle;

            %\draw [style=arrow] (8.center) to (9.center);

            \node [style=label, anchor=north] at (2.25, 0.75) {\small $\e(x_1)$};
            \node [style=label, anchor=north] at (0.55, 0.625) {\small $\e(x_2)$};
            \node [style=label, anchor=east] at (0.5, 2.5) {\small $\e(x_3)$};
            \node [style=point] at (4){};
            \node [style=point] at (5){};
            \node [style=point] at (6){};

            \node [style=label] at (8) {\small $A$};
        \end{pgfonlayer}
    \end{tikzpicture}
    \end{subfigure}
    \caption{some plots}
\end{figure}

\begin{frame}
    \frametitle{Wolff Cluster Algorithm}

    \begin{figure}[h]
      \centering
          \begin{subfigure}[b]{0.5\textwidth}\centering
            \includegraphics[width=0.6\textwidth]{imgs/wolffa.png}
            \caption{before cluster flip}
          \end{subfigure}%
          \hfill
          \begin{subfigure}[b]{0.5\textwidth}\centering
            \includegraphics[width=0.6\textwidth]{imgs/wolffb.png}
            \caption{after cluster flip}
          \end{subfigure}
          \hfill
          \caption{\label{fig:wolff} An example of the Wolff cluster algorithm in the $\phi^4$ model. White represents positive values of $\phi$ while black represents negative. $\lambda=0.5$, $m_0^2=-0.9$.}

    \end{figure}
\end{frame}

@@

@@

        %\begin{figure}[h]
  %\centering
      %\subfigure[$\tau=0$]{\includegraphics[width=0.20\textwidth]{imgs/0.png}}
      %\subfigure[$\tau=0.5$]{\includegraphics[width=0.20\textwidth]{imgs/0_5.png}}
      %\subfigure[$\tau=2$]{\includegraphics[width=0.20\textwidth]{imgs/2.png}} 
      %\subfigure[$\tau=10$]{\includegraphics[width=0.20\textwidth]{imgs/10.png}}
          %\caption{Effect of flow time evolution on a random lattice in the symmetric phase. Red and blue indicate positive and negative values of the field in 2D spacetime.}
  %\label{fig:flow}
%\end{figure}

@@

@@

    \end{enumerate}
    %\begin{figure}[h]

      %\centering
          %\subfigure[before]{\includegraphics[width=0.20\textwidth]{imgs/wolff1}}
          %\subfigure[after]{\includegraphics[width=0.20\textwidth]{imgs/wolff2}}
          %\caption{Wolff cluster algorithm applied on $\phi^4$ lattice in $1+1$ dimensions, red corresponding to positive values and green to negative}
      %\label{fig:flow}
    %\end{figure}
\end{frame}

@@

\begin{frame}
    \frametitle{Effect of $\theta$-term on $\langle Q \rangle$}
\begin{figure}[h]
  \centering
      \includegraphics[width=0.8\textwidth]{imgs/theta.png}
      %\caption{Absolute difference between Python and C++ results of average $\bar\phi$, Binder cumulant and susceptibility for various $m_0^2$ values. $\lambda=0.5$, $L=64$, $1000$ measurements, $1000$ sweep thermalization with measurements taken every $100$ sweeps. }
